Paper

### 2021-01-18 Paper.Rmd
...

@@

ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR + ggtitle(''), width=6, height=5, units='in', dpi=300)
rm(unadjRR, myPal)
```

@@

#### By pathology?
```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
pg.samp %>% dplyr::group_by(Status, Pathology, Risk) %>% tally(name='Sample count') %>% ungroup %>% dplyr::rename('Risk class' = 'Risk') %>%
  write_tsv('plots/discovery_cohort_by_path.tsv')

rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Discovery cohort samples', y='% samples classified', x='Sample pathology') + theme(strip.text = element_text(size=14))
rp
ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none') + ggtitle('')), width=7.5, height=6.5, units='in', dpi=300)
rm(rp)
```

@@

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
  geom_errorbar(color='grey55', width=0.2) +
  geom_text(aes(label=paste(round(pred.r*100), '%', sep=''), y=(pred.r-`sem(P)`)*100), color='white', nudge_y=-3) +
  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
  labs(x='Months prior to endpoint', y='Sensitivity for high risk samples', title='High risk classifications prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) ) 
rp

ggsave(filename='plots/pred_endo.png', plot=rp + ggtitle(''), width=6, height=5, units='in', dpi=300)
rm(endos,endos.m,rp)
```

@@

p = ggplot(risk.over.time) + facet_wrap(~Status, scales='free', ncol = 1) +
  geom_line(aes(x=months.since.initial, y=reorder(Patient, desc(ie.risk.diff)), color=ie.risk.diff)) + 
  scale_color_manual(values=unlist(riskColors()), name='Initial:Final Risk') + 
  geom_point(aes(x=months.since.initial, y=reorder(Patient, desc(ie.risk.diff)), fill=max.Risk, shape=max.Path), color='white', size=4 ) +  
  #geom_point(data=filter(risk.over.time, max.Path %in% c('HGD','IMC')), aes(x=months.since.initial, y=reorder(Patient, desc(ie.risk.diff))), shape=8, color='white', size=4 ) +  
  scale_fill_manual(values=unlist(riskColors()), name='Max Risk',guide=guide_legend(override.aes=list(fill=unlist(riskColors()), color=unlist(riskColors())))) + 
  scale_shape_manual(values=c(21,22,23,24,25), name='Max Pathology',guide=guide_legend(override.aes=list(fill='grey39', color='white'))) +
  labs(x='Months since initial endoscopy', y='Patient', title='Risks over time per endoscopy') + 
  plot.theme + theme(text = element_text(size=16))
p
ggsave(filename='plots/risks_over_time.png', plot = p + ggtitle(''), height=12, width=7, units='in')

@@

@@

rp = plot.risk.by.path(val.preds) + labs(title='Validation cohort samples', y='% samples classified', x='Sample pathology')  + theme(strip.text = element_text(size=12))
rp
ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp + theme(legend.position = 'none') + ggtitle(''), width=6, height=7, units='in', dpi=300)
rm(rp)

@@

p = second %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% dplyr::rename(RFA = `1`, `6-12 month recheck` = `2`, `12-24 month recheck` = `3`, `3-5 year surveillance` = `4`)  %>%
  melt(id.vars = 'Status') %>% mutate(Status = recode(Status, NP = 'Non-Progressor' ,P = 'Progressor')) %>%
  ggplot(aes(variable, value, group = Status)) + 
  geom_col(aes(fill=Status), position = position_dodge()) + geom_text(aes(label=paste0('n=',value)), position = position_dodge(width = 1)) + coord_flip() +
  scale_fill_manual(values = c('#4575B4', '#A50026'), name='' ) + plot.theme + labs(x='', y='Number of Patients', title='Second Endoscopy Rx') + theme(legend.position = 'bottom')
p
ggsave('plots/second_endo_rx.png', plot=p, width=6, height=6, units='in', dpi=300)
rm(p)

>>>>

globalCNdf %>% filter(seg.type != 'arms') %>% dplyr::select(Patient, Status, Pathology, variable, chr, start, end, value) %>% 
  dplyr::rename_all(list(~str_to_title(.))) %>% dplyr::rename('Chromosome' = 'Chr', 'Sample' = 'Variable', 'CN Value' = 'Value') %>%
  write_tsv('plots/global_cn_values.tsv')

@@

ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR + labs(title='', x='', y=''), width=6, height=5, units='in', dpi=300)
rm(unadjRR, myPal)
```

@@

#### By pathology?
```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
#tmp = pg.samp %>% dplyr::mutate(Pathology = recode(Pathology, 'HGD' = 'HGD/IMC', 'IMC' = 'HGD/IMC'))
pg.samp %>% dplyr::group_by(Status, Pathology, Risk) %>% tally(name='Sample count') %>% ungroup %>% dplyr::rename('Risk class' = 'Risk') %>%
  write_tsv('plots/discovery_cohort_by_path.tsv')

rp = plot.risk.by.path(pg.samp %>% ungroup, label_bars = F, ret_df = F)

rp + labs(title='Discovery cohort samples', y='% samples classified', x='Sample pathology') + theme(strip.text = element_text(size=14))

ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none') + labs(title='', x='',y='')), width=7.5, height=6.5, units='in', dpi=300)
rm(rp)
```

@@

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
  geom_errorbar(color='grey55', width=0.2) +
  labs(x='Months prior to endpoint', y='Sensitivity for high risk samples', title='High risk classifications prior to final endoscopy') + coord_cartesian(ylim = c(0,100)) + 
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) )

rp + geom_text(aes(label=paste(round(pred.r*100), '%', sep=''), y=(pred.r-`sem(P)`)*100), color='white', nudge_y=-3) +
  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + labs(subtitle='n = #endoscopies, s = #samples')

ggsave(filename='plots/pred_endo.png', plot=rp + labs(title='', x='', y=''), width=6, height=5, units='in', dpi=300)
rm(endos,endos.m,rp)
```

@@

p = ggplot(risk.over.time) + facet_wrap(~Status, scales='free', ncol = 2) +
  geom_line(aes(x=months.since.initial, y=reorder(Patient, desc(ie.risk.diff)), color=ie.risk.diff)) + 
  scale_color_manual(values=unlist(riskColors()), name='Initial:Final Risk') + 
  geom_point(aes(x=months.since.initial, y=reorder(Patient, desc(ie.risk.diff)), fill=max.Risk, shape=max.Path), color='white', size=4 ) +  
  #geom_point(data=filter(risk.over.time, max.Path %in% c('HGD','IMC')), aes(x=months.since.initial, y=reorder(Patient, desc(ie.risk.diff))), shape=8, color='white', size=4 ) +  
  scale_fill_manual(values=unlist(riskColors()), name='Max Risk',guide=guide_legend(override.aes=list(fill=unlist(riskColors()), color=unlist(riskColors())))) + 
  scale_shape_manual(values=c(21,22,23,24,25), name='Max Pathology',guide=guide_legend(override.aes=list(fill='grey39', color='white'))) +
  labs(x='Months since initial endoscopy', y='Patient', title='Risks over time per endoscopy') + 
  plot.theme + theme(text=element_text(size=16), panel.spacing.x=unit(1,'lines')  )
p

legend = get.legend(p)

ggsave(filename='plots/risks_over_time.png', plot = p + labs(title='',x='',y='') + theme(legend.position='none'), height=8, width=12, units='in')

ggsave(filename='plots/risks_over_time_legend.png', plot = grid.arrange(legend), height=5, width=2, units='in')

@@

# How predictive are the strongest coefs?
```{r}
features = features %>% arrange(desc(hazards))

ftsamples = tibble(); ftpatients = tibble()
for (i in 1:nrow(features)) {
  ft = features %>% dplyr::slice(i)
  ft_cn = globalCNdf %>% filter(chr == ft$chr & start >= ft$start & end <= ft$end & 
                      seg.type == case_when((ft$end-ft$start) > 5e6 ~ 'arms', TRUE ~ '5MB') &
                      as.character(cn) == as.character(ft$gl))

  ftsamples = bind_rows(ftsamples, ft_cn %>% dplyr::group_by(Status) %>% tally() %>% spread(Status, n) %>% 
    dplyr::mutate('Sample counts for:' = paste0(ft$gl, ' of ', ft$label)) %>% 
    dplyr::select('Sample counts for:', everything()))

  ftpatients = bind_rows(ftpatients, ft_cn %>% dplyr::select(Patient, Status) %>% distinct %>% 
    group_by(Status) %>%  tally() %>% spread(Status, n) %>%
    dplyr::mutate('Patient counts for:' = paste0(ft$gl, ' of ', ft$label)) %>% 
    dplyr::select('Patient counts for:', everything()))
}

ftsamples = ftsamples %>% dplyr::mutate_all(list(~replace_na(.,0)))
ftpatients = ftpatients %>% dplyr::mutate_all(list(~replace_na(.,0)))

ftpatients %>% arrange(desc(Progressor))
ftsamples %>% arrange(desc(Progressor))

@@

plot.risk.by.path(val.preds, label_bars = T) + labs(title='Validation cohort samples', y='% samples classified', x='Sample pathology')  + theme(strip.text = element_text(size=12))

rp = plot.risk.by.path(val.preds, label_bars = F) + labs(title='Validation cohort samples', y='% samples classified', x='Sample pathology')  + theme(strip.text = element_text(size=12))
rp
ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp + theme(legend.position = 'none') + labs(title='',x='',y=''), width=6, height=7, units='in', dpi=300)
rm(rp)

@@

p = second %>% group_by(Status, Rule) %>% tally %>% spread(Rule,n) %>% dplyr::rename(RFA = `1`, `6-12 month recheck` = `2`, `12-24 month recheck` = `3`, `3-5 year surveillance` = `4`)  %>%
  melt(id.vars = 'Status') %>% mutate(Status = recode(Status, NP = 'Non-Progressor' ,P = 'Progressor')) %>%
  ggplot(aes(variable, value, group = Status)) + scale_y_continuous(expand=c(0,0), limits=c(0,25)) +
  geom_col(aes(fill=Status), position = position_dodge()) + 
  coord_flip() +
  scale_fill_manual(values = c('#4575B4', '#A50026'), name='' ) + plot.theme + 
  labs(x='', y='Number of Patients', title='Second Endoscopy Rx') + theme(legend.position = 'bottom')
p + geom_text(aes(label=paste0('n=',value)), position = position_dodge(width = 1))

ggsave('plots/second_endo_rx.png', plot=p, width=6, height=6, units='in', dpi=300)
rm(p)
